---
title: "Electricity/Solar Update"
output: html_notebook
---

A look at the latest Jedlovec House eletricity usage and solar production from PPL and Enphase.

Load packages
```{r}
library(tidyverse)
library(lubridate)
library(hms)
library(readxl)
library(onlineBcp)
```

```{r}
hourly_ppl_pivot <- read_csv("ppl_15mins.csv")

(hourly_ppl_net <- hourly_ppl_pivot %>% 
  filter(`Read Type` == "kWh Net"))

```

Load Enphase data
```{r}
full_hist <- read_csv("enphase_production.csv")

(hourly_production <- full_hist %>% 
  rename(datetime = `Date/Time`, energy_produced_Wh = `Energy Produced (Wh)`) %>% 
  mutate(datetime = as.POSIXct(datetime,format="%m/%d/%Y %H:%M",tz=Sys.timezone())) %>% 
  mutate(date = date(datetime), time = as.hms(format(datetime, format = "%H:%M:%S")), month = month(datetime, label=TRUE), year = year(datetime), day = day(datetime), yday = yday(datetime), monthday = format(datetime, "%m-%d"), wday = wday(datetime, label=TRUE), equinox_day = (yday + 10) %% 365, equinox_group = floor((equinox_day+15)/30)*30)
)

```

Spot-check Enphase
Should see lifetime production by day and by hour
```{r}
ggplot(hourly_production, aes(datetime, energy_produced_Wh)) +
  geom_point()

ggplot(hourly_production, aes(time, energy_produced_Wh)) +
  theme(axis.text.x = element_text(angle = 90)) +
  geom_point()

```

Net + Produced = Consumed

```{r}
# hourly_ppl_net <- hourly_ppl_net %>% mutate(date = as_date(date))

#hourly_ppl_net %>% arrange(desc(date))
#hourly_production %>% arrange(desc(date))

(hourly_electricity <- hourly_ppl_net %>% 
    inner_join(hourly_production, by = join_by(date,time))  %>% 
    mutate(consumed_kWh = kWh + energy_produced_Wh/1000, produced_kWh = energy_produced_Wh/1000)  %>% 
    rename(net_kWh = kWh) %>% 
    select(datetime, date, time, net_kWh, produced_kWh, consumed_kWh) %>%
    arrange(desc(datetime)))

```

Calculate production for first year of solar panels, second year, etc.
```{r}

hourly_electricity <- hourly_electricity %>% 
  mutate(solar_year = as_factor(ceiling(interval(as_date(as.POSIXct('04/15/2022',format="%m/%d/%Y",tz=Sys.timezone())),as_date(date)) / years(1) ))) %>% 
  mutate(yday = yday(date)) %>% 
  arrange(datetime)

hourly_electricity %>% 
  group_by(solar_year) %>% 
  filter(solar_year == 1 | solar_year == 2) %>% 
  summarize(net_kWh = sum(net_kWh), produced_kWh = sum(produced_kWh), consumed_kWh = sum(consumed_kWh)) 

```

Interesting! Produced kWh went down by 5%, but consumed kWh went down by 10% despite the fact that we got an electric car! Let's explore that further. 

```{r}
yday_order <- c(106:366,1:105)

daily_electricity <- hourly_electricity %>% 
  filter(solar_year == 1 | solar_year == 2) %>% 
  group_by(solar_year, yday) %>% 
  summarize(net_kWh = sum(net_kWh), produced_kWh = sum(produced_kWh), consumed_kWh = sum(consumed_kWh), .groups = "drop") %>% 
  group_by(neworder = cumsum(yday == 106)) %>% 
  mutate(plot_date = factor(yday, levels = yday_order, ordered = TRUE)) %>%
  #mutate(plot_date = factor(yday, levels = yday_order, labels = parse_date_time(yday, orders = "j"), ordered = TRUE)) %>%
  arrange(plot_date)

ggplot(daily_electricity, aes(plot_date, consumed_kWh, group=solar_year, color=solar_year)) + 
  geom_point() +
  geom_smooth(span=0.3) +
  #scale_x_continuous(breaks = c(1, 32, 60, 91, 121, 152, 182, 213, 244, 274, 305, 335), labels = month.abb) +
  theme(axis.text.x = element_text(angle = 90)) 

```

Daily totals
```{r}

daily_electricity <- hourly_electricity %>% 
  filter(solar_year == 1 | solar_year == 2) %>% 
  mutate(solar_year_start_year = as.numeric(solar_year) - 1 + 2021) %>% 
  mutate(solar_year_start_date = paste('04/15/',solar_year_start_year,sep='')) %>% 
  mutate(solar_day = interval(as_date(as.POSIXct(paste('04/15/',solar_year_start_year),format="%m/%d/%Y",tz=Sys.timezone())),as_date(date)) / days(1) 
         ) %>% 
  #filter(solar_day < 367) %>% 
  group_by(date, solar_day, solar_year) %>% 
  summarize(net_kWh = sum(net_kWh), produced_kWh = sum(produced_kWh), consumed_kWh = sum(consumed_kWh))

ggplot(daily_electricity, aes(solar_day, consumed_kWh, group = solar_year, color = solar_year)) + 
  geom_point() +
  #scale_x_continuous(breaks = c(1, 32, 60, 91, 121, 152, 182, 213, 244, 274, 305, 335), labels = month.abb) +
  theme(axis.text.x = element_text(angle = 90)) 

 
```
Look how our electricity consumption got much less predictable, more uneven, after we purchased an EV and installed a level 2 charger in late January/early February 2023!

Also, a couple of outlier high points in Dec 2022 before we got the EV... hmm.

We probably need to filter out the EV charging and compare that separately. 

Average consumption based on time of day
```{r}
#hourly_electricity %>% summarize(min_date = min(date), max_date = max(date))

electricity_by_time <- hourly_electricity %>% 
  filter(solar_year == 1 | solar_year == 2) %>% 
  #filter(date <= as.POSIXct('12/31/2022 00:00',format="%m/%d/%Y %H:%M",tz=Sys.timezone()) ) %>% 
  group_by(time, solar_year) %>% 
  summarize(produced_kWh = mean(produced_kWh), consumed_kWh = mean(consumed_kWh), net_kWh = mean(net_kWh)) %>% 
  arrange(time)

electricity_by_time

ggplot(electricity_by_time, aes(x=time, y=consumed_kWh, group=solar_year, color=solar_year)) +
  geom_point() 
  #labs(colour="",x="Time of Day",y="Electricity Consumption (kWh)")+
  #scale_color_manual(values = c("red","black","green")) +
  #ggtitle("Home Electricity in 15-Minute Intervals (Since April 15, 2022)")

```
Wow, look how much less was used during the day in year 2!!! Like a 50% reduction! 

Is this due to milder weather? 

Let's look at it by month: 

```{r}
#electricity_by_month <- 
hourly_electricity %>% 
  mutate(month = month(date)) %>% 
  group_by(solar_year, month) %>% 
  summarize(consumed_kWh = mean(consumed_kWh)) %>% 
  arrange(month, solar_year) %>% 
  pivot_wider(names_from = month, values_from = consumed_kWh) %>% 
  arrange(solar_year)

# hourly_electricity %>% 
#   group_by(solar_day) %>% 
#   summarize(consumed_kWh = mean(consumed_kWh)) %>% 
#   arrange(month, solar_year) %>% 
#   pivot_wider(names_from = month, values_from = consumed_kWh) %>% 
#   arrange(solar_year)



#Look at limited window during day (no EV charging)
 
# hourly_electricity %>% 
#   filter(time == '12:00:00') %>% 
#   #filter(time > as.POSIXct('04:00:00',format="%H:%M:$s",tz=Sys.timezone()) ) %>%  #& time < as.POSIXct('18:00',format="%H:%M",tz=Sys.timezone())) %>% 
#   mutate(month = month(date)) %>% 
#   group_by(solar_year, month) %>% 
#   summarize(consumed_kWh = mean(consumed_kWh)) %>% 
#   arrange(month, solar_year) %>% 
#   pivot_wider(names_from = month, values_from = consumed_kWh) %>% 
#   arrange(solar_year)


```

Look at specific months

```{r}
electricity_by_month_time <- hourly_electricity %>% 
  mutate(month = as_factor(month(date))) %>% 
  filter(solar_year == 1 | solar_year == 2) %>% 
  #Sample month
  filter(month == 6) %>% 
  group_by(solar_year, month, time) %>% 
  summarize(produced_kWh = mean(produced_kWh), consumed_kWh = mean(consumed_kWh), net_kWh = mean(net_kWh)) %>% 
  arrange(solar_year, month, time)

ggplot(electricity_by_month_time, aes(time, consumed_kWh, group=solar_year, color=solar_year)) +
  #facet_grid(rows = vars(month)) +
  geom_point(size=0.5) +
  ggtitle("Home Electricity in 15-Minute Intervals") +
  ylab("Electricity Consumption (kWh)")

```

Example day with EV charging

```{r}
sample_day <- hourly_electricity %>% 
  filter(solar_year == 1 | solar_year == 2) %>% 
  filter(yday == 161) %>% 
  group_by(date, solar_year, yday, time) %>% 
  summarize(produced_kWh = mean(produced_kWh), consumed_kWh = mean(consumed_kWh), net_kWh = mean(net_kWh)) %>% 
  arrange(date, solar_year, yday, time)

ggplot(sample_day, aes(time, consumed_kWh, group=solar_year, color=solar_year)) +
  #facet_grid(rows = vars(month)) +
  geom_point(size=0.5) +
  ggtitle("Home Electricity in 15-Minute Intervals") +
  ylab("Electricity Consumption (kWh)")

```

Ok, let's detect and remove EV charging sessions. 

Test on electricity consumption data
```{r}
# library(onlineBcp)

#Select a sample to test
sample <- hourly_electricity %>% 
  mutate(solar_day = interval(as_date(as.POSIXct('04/15/2022',format="%m/%d/%Y",tz=Sys.timezone())),as_date(date)) / days(1) ) %>%
  filter(solar_year == 2) %>% 
  filter(yday < 38 & yday > 35)

ggplot(sample, aes(datetime, consumed_kWh, group=solar_year, color=solar_year)) +
  #facet_grid(rows = vars(month)) +
  geom_point(size=0.5) +
  ggtitle("Home Electricity in 15-Minute Intervals") +
  ylab("Electricity Consumption (kWh)")
```

bcp model comparisons: https://lindeloev.github.io/mcp/articles/packages.html


```{r}
min_time <- pull(sample %>% summarize(min_time = min(datetime)))

sample <- sample %>% mutate(time_x = interval(min_time,datetime)/minutes(15)+1)

#Filter data to feed to model
x <- sample %>%  
  ungroup() %>% 
  select(consumed_kWh)

x <- x$consumed_kWh

#Run online Bayesian Change Point model 
#"Online" means in progress, not retroactive
bcp <- online_cp(x)

plot(summary(bcp))

```
Not too bad! It picks up the EV charging sessions, but it also picks up some other, presumably HVAC-related consumption changes. It looks like I should be able to weed these out easily. 


```{r}

#Select a sample to test
sample <- hourly_electricity %>% 
  #filter(solar_year == 2) %>% 
  mutate(solar_day = interval(as_date(as.POSIXct('04/15/2022',format="%m/%d/%Y",tz=Sys.timezone())),as_date(date)) / days(1) ) 

ggplot(sample, aes(datetime, consumed_kWh, group=solar_year, color=solar_year)) +
  #facet_grid(rows = vars(month)) +
  geom_point(size=0.5) +
  ggtitle("Home Electricity in 15-Minute Intervals") +
  ylab("Electricity Consumption (kWh)")
```
You can pick out our family vacation in August, and you can also see a period of high electricity usage in February, presumably related to cold weather. 
- Verify with weather data? 


```{r}
#sample %>% arrange(datetime)

min_time <- pull(sample %>% summarize(min_time = min(datetime)))

sample <- sample %>% 
  mutate(time_x = interval(min_time,datetime)/minutes(15)+1) %>% 
  arrange(datetime)

#Filter data to feed to model
x <- sample %>%  
  ungroup() %>% 
  select(consumed_kWh)

x <- x$consumed_kWh

#Run online Bayesian Change Point model 
#"Online" means in progress, not retroactive
bcp <- online_cp(x)

summary(bcp)

plot(summary(bcp))
```
Nice! Looks like it still works over a longer period. (An advantage of an onlineBcp model?)

It looks like any interval with a mean above about 2.0 kWh should qualify as an EV charging session. 
- Do we need to do anything to avoid mixups with high HVAC usage? 
- Do we need to put an ideal length of charge on the intervals? Maximum variation during the period? 
- Use post.prob to filter out any intervals? 

```{r}
result <- summary(bcp, norm.test = TRUE)

bcp_segments <- as_tibble(result$result$segment) %>% 
  mutate(length = end - begin) %>% 
  mutate(lag_mean = lag(mean,1)) %>% 
  mutate(lead_mean = lead(mean,1)) %>% 
  mutate(adj = (lag_mean+lead_mean)/2)

bcp_segments %>% 
  filter(mean > 2.0 & mean < 3.0) 

```

```{r}

sample_adj <-  sample %>%  
  left_join(bcp_segments, by = join_by(time_x>=begin,time_x<=end)) %>% 
  mutate(consumed_kWh_adj = case_when(
                                mean > 2.0 & length >= 4 & SD < 1.1 ~ consumed_kWh - (mean - adj), 
                                .default = consumed_kWh
                                )
         ) %>% 
  mutate(consumed_kWh_adj = case_when(
                                consumed_kWh_adj < 0 ~ 0, 
                                .default = consumed_kWh_adj
                                )
  )

sample_adj %>% 
  mutate(diff = consumed_kWh - consumed_kWh_adj) %>% 
  group_by(solar_year) %>% 
  summarize(diff = sum(diff))

ggplot(sample_adj, aes(datetime, consumed_kWh_adj, group=solar_year, color=solar_year)) +
  #facet_grid(rows = vars(month)) +
  geom_point(size=0.5) +
  ggtitle("Home Electricity in 15-Minute Intervals") +
  ylab("Electricity Consumption (kWh)")

```


solar_year
<fctr>
diff
<dbl>
0	-0.3670			
1	595.8770			
2	3504.1387			
3	321.9421			

vs. 

solar_year
<fctr>
kWh
<dbl>
1	683			
2	4382			
3	607

Underestimates  

Ignore the bcp model, just pull data on home charging from Emporia

```{r}
emporia_daily <- read_csv('/Users/benjedlovec/IoniqCharging/emporia_daily.csv') %>% 
  mutate(date = date(datetime))

daily_electricity_adj <- daily_electricity %>% 
  mutate(date = as_date(date)) %>% 
  left_join(emporia_daily, by = join_by(date)) %>% 
  mutate(consumed_kWh_adj = consumed_kWh - replace_na(kWh,0)) %>% 
  rename(solar_year = solar_year.x, charging_kWh = kWh)

daily_electricity_adj %>% 
  mutate(diff = consumed_kWh - consumed_kWh_adj) %>% 
  group_by(solar_year) %>% 
  summarize(consumed_kWh = sum(consumed_kWh), charging_kWh = sum(replace_na(charging_kWh,0)), consumed_kWh_adj = sum(replace_na(consumed_kWh_adj,0)), diff = sum(replace_na(diff,0)))

ggplot(daily_electricity_adj, aes(solar_day, consumed_kWh_adj, group = solar_year, color = solar_year)) + 
  geom_point() +
  #scale_x_continuous(breaks = c(1, 32, 60, 91, 121, 152, 182, 213, 244, 274, 305, 335), labels = month.abb) +
  theme(axis.text.x = element_text(angle = 90)) 


```

```{r}
monthly_reduction <- daily_electricity_adj %>% 
  mutate(year_1 = sum(case_when(solar_year == 1 ~ consumed_kWh_adj)), year_2 = sum(case_when(solar_year == 2 ~ consumed_kWh_adj))) %>% 
  group_by(month = month(date)) %>% 
  summarize(year_1 = round(sum(replace_na(year_1,0))), year_2 = round(sum(replace_na(year_2,0)))) %>% 
  mutate(pct = round((year_1-year_2)/year_1,2))

monthly_reduction

daily_electricity_adj %>% 
  group_by(month = month(date), solar_year) %>% 
  summarize(consumed_kWh_adj = sum(consumed_kWh_adj))

```

