---
title: "Home Electricity Emissions Data"
output: html_notebook
---

Use tibbles created in "WattTimeMOERData" and "Hourly Electricity Consumption" notebooks
```{r}
hourly_electricity 
moer

hourly_electricity %>% 
  summarize(min_time = min(datetime), max_time = max(datetime))

moer %>% 
  summarize(min_time = min(timestamp), max_time = max(timestamp))

```

NOTE: MOER data from CAISO_NORTH (free) from WattTime as proxy for PJM data

Check time zone alignment - believe that everything is in local time

Join tables

MOER is lbs/MWh
Convert to kg/kWh? 
1 lbs = 0.45 kg
```{r}
hourly_emissions <- hourly_electricity %>% 
  inner_join(moer, by = join_by(datetime == local_timestamp)) %>% 
  select(datetime, net_kWh, produced_kWh, consumed_kWh, MOER) %>% 
  mutate(moer_kg_kWh = MOER*0.45/1000) %>% 
  mutate(net_emissions_co2_kg = moer_kg_kWh*net_kWh, produced_emissions_co2_kg = moer_kg_kWh*produced_kWh, consumed_emissions_co2_kg = moer_kg_kWh*consumed_kWh) %>% 
  arrange(datetime)

hourly_emissions %>% 
  summarize(min_time = min(datetime), max_time = max(datetime))

hourly_emissions
```

Net energy and emissions over this time period
```{r}

hourly_emissions %>% 
  summarize(net_kWh = sum(net_kWh), produced_kWh = sum(produced_kWh), consumed_kWh = sum(consumed_kWh), net_emissions_co2_kg = sum(net_emissions_co2_kg), produced_emissions_co2_kg = sum(produced_emissions_co2_kg), consumed_emissions_co2_kg = sum(consumed_emissions_co2_kg)) %>% 
  mutate(pct_reduction = 1 - (consumed_emissions_co2_kg - produced_emissions_co2_kg)/consumed_emissions_co2_kg)

```

```{r}

emissions_date <- hourly_emissions %>% 
  mutate(date = date(datetime)) %>% 
  group_by(date) %>% 
  summarize(net_kWh = sum(net_kWh), produced_kWh = sum(produced_kWh), consumed_kWh = sum(consumed_kWh), net_emissions_co2_kg = sum(net_emissions_co2_kg), produced_emissions_co2_kg = sum(produced_emissions_co2_kg), consumed_emissions_co2_kg = sum(consumed_emissions_co2_kg)) %>% 
  mutate(pct_reduction = 1 - (consumed_emissions_co2_kg - produced_emissions_co2_kg)/consumed_emissions_co2_kg)

emissions_date

ggplot(emissions_date, aes(date,net_emissions_co2_kg)) +
  geom_point() 

```

Need to fix monthyear

```{r}

emissions_month <- hourly_emissions %>% 
  mutate(monthyear = paste(year(datetime),"-",months(datetime), sep = "")) %>% 
  group_by(monthyear) %>% 
  summarize(net_kWh = sum(net_kWh), produced_kWh = sum(produced_kWh), consumed_kWh = sum(consumed_kWh), net_emissions_co2_kg = sum(net_emissions_co2_kg), produced_emissions_co2_kg = sum(produced_emissions_co2_kg), consumed_emissions_co2_kg = sum(consumed_emissions_co2_kg)) %>% 
  mutate(pct_reduction = 1 - (consumed_emissions_co2_kg - produced_emissions_co2_kg)/consumed_emissions_co2_kg)

emissions_month

ggplot(emissions_month, aes(monthyear,net_emissions_co2_kg)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1)) 
```


```{r}

hourly_emissions_365 <- hourly_emissions %>% 
  filter(datetime >= as.POSIXct('11/01/2022 00:00',format="%m/%d/%Y %H:%M",tz=Sys.timezone()) & datetime < as.POSIXct('11/01/2023 00:00',format="%m/%d/%Y %H:%M",tz=Sys.timezone()))


hourly_emissions_365 %>% 
  summarize(days = n()/24/4, net_kWh = sum(net_kWh), produced_kWh = sum(produced_kWh), consumed_kWh = sum(consumed_kWh), net_emissions_co2_kg = sum(net_emissions_co2_kg), produced_emissions_co2_kg = sum(produced_emissions_co2_kg), consumed_emissions_co2_kg = sum(consumed_emissions_co2_kg)) %>% 
  mutate(pct_reduction = 1 - (consumed_emissions_co2_kg - produced_emissions_co2_kg)/consumed_emissions_co2_kg)


```
74% emissions reduction with solar panels based on CAISO NORTH MOER data
- Would expect higher with PJM

Wow, I was expecting consumption to exceed production since the electric car. 
- Did efficiency gains (HVAC?) offset the car charging? 
- Compare month 2022 (May, June, etc.) to month 2023 by time of day for consumption
- Perhaps used less during the day and more when car charging at night? 
- Temperature related (less HVAC required this year than last?)


