---
title: "Electricity/Solar Update"
output: html_notebook
---

A look at the latest Jedlovec House eletricity usage and solar production from PPL and Enphase.

Load packages
```{r}
library(tidyverse)
library(lubridate)
library(hms)
library(readxl)
library(onlineBcp)
```

Load PPL data
```{r}

col_datatypes <- c('numeric','numeric','date','text',rep('numeric',99))

hourly1 <- read_excel("Hourly Usage 220326 to 220624.xlsx", col_types = col_datatypes)
hourly2 <- read_excel("Hourly Usage 220625 to 230623.xlsx", col_types = col_datatypes)
hourly3 <- read_excel("PPL 230624 to 240509.xlsx", col_types = col_datatypes)

#hourly1
#hourly2
#hourly3

ppl_15mins <- bind_rows(hourly1,list(hourly2,hourly3))

ppl_15mins %>% arrange(desc(Date), `Read Type`)


```

Transform PPL Data
```{r}

hourly_ppl_pivot <- ppl_15mins %>% 
  rename(date = Date) %>% 
  pivot_longer(!c("Account Number", "Meter Number", date, "Read Type", Min, Max, Total), names_to = "time", values_to = "kWh") 

rename(ppl_15mins, date = Date)

hourly_ppl_pivot <- hourly_ppl_pivot %>% 
  mutate(time = parse_time(time, '%H:%M %p'), month = month(date, label=TRUE), year = year(date), yday = yday(date), wday = wday(date, label=TRUE))

(hourly_ppl_net <- hourly_ppl_pivot %>% 
  filter(`Read Type` == "kWh Net"))
```

Load Enphase data
```{r}
import <- read_csv("enphase_history_230701.csv")
update <- read_csv("hourly_generation_230625_240510.csv")

update <- update %>% 
  mutate(`Date/Time` = as.POSIXct(`Date/Time`,format="%m/%d/%Y %H:%M",tz=Sys.timezone())) %>% 
  filter(`Date/Time` >= as.POSIXct('07/01/2023 00:00',format="%m/%d/%Y %H:%M",tz=Sys.timezone()) )

import <- import %>% 
  mutate(`Date/Time` = as.POSIXct(`Date/Time`,format="%m/%d/%Y %H:%M",tz=Sys.timezone())) 

full_hist <- rbind(import,update)

full_hist %>% arrange(desc(`Date/Time`))

(hourly_production <- full_hist %>% 
  rename(datetime = `Date/Time`, energy_produced_Wh = `Energy Produced (Wh)`) %>% 
  mutate(datetime = as.POSIXct(datetime,format="%m/%d/%Y %H:%M",tz=Sys.timezone())) %>% 
  mutate(date = date(datetime), time = as.hms(format(datetime, format = "%H:%M:%S")), month = month(datetime, label=TRUE), year = year(datetime), day = day(datetime), yday = yday(datetime), monthday = format(datetime, "%m-%d"), wday = wday(datetime, label=TRUE), equinox_day = (yday + 10) %% 365, equinox_group = floor((equinox_day+15)/30)*30)
)

```

Spot-check Enphase
Should see lifetime production by day and by hour
```{r}
ggplot(hourly_production, aes(datetime, energy_produced_Wh)) +
  geom_point()

ggplot(hourly_production, aes(time, energy_produced_Wh)) +
  theme(axis.text.x = element_text(angle = 90)) +
  geom_point()

```

Net + Produced = Consumed

```{r}
# hourly_ppl_net <- hourly_ppl_net %>% mutate(date = as_date(date))

#hourly_ppl_net %>% arrange(desc(date))
#hourly_production %>% arrange(desc(date))

(hourly_electricity <- hourly_ppl_net %>% 
    inner_join(hourly_production, by = join_by(date,time))  %>% 
    mutate(consumed_kWh = kWh + energy_produced_Wh/1000, produced_kWh = energy_produced_Wh/1000)  %>% 
    rename(net_kWh = kWh) %>% 
    select(datetime, date, time, net_kWh, produced_kWh, consumed_kWh) %>%
    arrange(date))

```

Calculate production for first year of solar panels, second year, etc.

```{r}

hourly_electricity <- hourly_electricity %>% 
  mutate(solar_year = factor(case_when(
                        datetime < as.POSIXct('04/16/2022 00:00',format="%m/%d/%Y %H:%M",tz=Sys.timezone()) ~ 0,
                        datetime >= as.POSIXct('04/16/2022 00:00',format="%m/%d/%Y %H:%M",tz=Sys.timezone())  &
                          datetime < as.POSIXct('04/16/2023 00:00',format="%m/%d/%Y %H:%M",tz=Sys.timezone()) ~ 1,
                        datetime >= as.POSIXct('04/16/2023 00:00',format="%m/%d/%Y %H:%M",tz=Sys.timezone())  &
                          datetime < as.POSIXct('04/16/2024 00:00',format="%m/%d/%Y %H:%M",tz=Sys.timezone()) ~ 2,
                        TRUE ~ 3)
         )) %>% 
  mutate(yday = yday(date))

hourly_electricity %>% 
  group_by(solar_year) %>% 
  summarize(net_kWh = sum(net_kWh), produced_kWh = sum(produced_kWh), consumed_kWh = sum(consumed_kWh))

```

Interesting! Produced kWh went down by 5%, but consumed kWh went down by 10% despite the fact that we got an electric car! Let's explore that further. 

```{r}
daily_electricity <- hourly_electricity %>% 
  group_by(solar_year, yday) %>% 
  summarize(net_kWh = sum(net_kWh), produced_kWh = sum(produced_kWh), consumed_kWh = sum(consumed_kWh))

ggplot(daily_electricity, aes(yday, consumed_kWh, group=solar_year, color=solar_year)) + 
  geom_point() +
  geom_smooth(span=0.3) +
  scale_x_continuous(breaks = c(1, 32, 60, 91, 121, 152, 182, 213, 244, 274, 305, 335), labels = month.abb) +
  theme(axis.text.x = element_text(angle = 90)) 

```

Daily totals
```{r}
daily_electricity <- hourly_electricity %>% 
  mutate(solar_day = interval(as_date(as.POSIXct('04/15/2022',format="%m/%d/%Y",tz=Sys.timezone())),as_date(date)) / days(1) 
         ) %>% 
  group_by(date, solar_day) %>% 
  summarize(net_kWh = sum(net_kWh), produced_kWh = sum(produced_kWh), consumed_kWh = sum(consumed_kWh))

ggplot(daily_electricity, aes(date, consumed_kWh)) + 
  geom_point() +
  #scale_x_continuous(breaks = c(1, 32, 60, 91, 121, 152, 182, 213, 244, 274, 305, 335), labels = month.abb) +
  theme(axis.text.x = element_text(angle = 90)) 

 
```
Look how our electricity consumption got much less predictable, more uneven, after we purchased an EV and installed a level 2 charger in late January/early February 2023!

Also, a couple of outlier high points in Dec 2022 before we got the EV... hmm.

We probably need to filter out the EV charging and compare that separately. 

Average consumption based on time of day
```{r}

#hourly_electricity %>% summarize(min_date = min(date), max_date = max(date))

electricity_by_time <- hourly_electricity %>% 
  filter(solar_year == 1 | solar_year == 2) %>% 
  #filter(date <= as.POSIXct('12/31/2022 00:00',format="%m/%d/%Y %H:%M",tz=Sys.timezone()) ) %>% 
  group_by(time, solar_year) %>% 
  summarize(produced_kWh = mean(produced_kWh), consumed_kWh = mean(consumed_kWh), net_kWh = mean(net_kWh)) %>% 
  arrange(time)

electricity_by_time

ggplot(electricity_by_time, aes(x=time, y=consumed_kWh, group=solar_year, color=solar_year)) +
  geom_point() 
  #labs(colour="",x="Time of Day",y="Electricity Consumption (kWh)")+
  #scale_color_manual(values = c("red","black","green")) +
  #ggtitle("Home Electricity in 15-Minute Intervals (Since April 15, 2022)")

```
Wow, look how much less was used during the day in year 2!!! Like a 50% reduction! 

Is this due to milder weather? 

Let's look at it by month: 

```{r}
#electricity_by_month <- 
hourly_electricity %>% 
  mutate(month = month(date)) %>% 
  group_by(solar_year, month) %>% 
  summarize(consumed_kWh = mean(consumed_kWh)) %>% 
  arrange(month, solar_year) %>% 
  pivot_wider(names_from = month, values_from = consumed_kWh) %>% 
  arrange(solar_year)

#Look at limited window during day (no EV charging)
 
# hourly_electricity %>% 
#   filter(time == '12:00:00') %>% 
#   #filter(time > as.POSIXct('04:00:00',format="%H:%M:$s",tz=Sys.timezone()) ) %>%  #& time < as.POSIXct('18:00',format="%H:%M",tz=Sys.timezone())) %>% 
#   mutate(month = month(date)) %>% 
#   group_by(solar_year, month) %>% 
#   summarize(consumed_kWh = mean(consumed_kWh)) %>% 
#   arrange(month, solar_year) %>% 
#   pivot_wider(names_from = month, values_from = consumed_kWh) %>% 
#   arrange(solar_year)


```

Look at specific months

```{r}
electricity_by_month_time <- hourly_electricity %>% 
  mutate(month = as_factor(month(date))) %>% 
  filter(solar_year == 1 | solar_year == 2) %>% 
  #Sample month
  filter(month == 6) %>% 
  group_by(solar_year, month, time) %>% 
  summarize(produced_kWh = mean(produced_kWh), consumed_kWh = mean(consumed_kWh), net_kWh = mean(net_kWh)) %>% 
  arrange(solar_year, month, time)

ggplot(electricity_by_month_time, aes(time, consumed_kWh, group=solar_year, color=solar_year)) +
  #facet_grid(rows = vars(month)) +
  geom_point(size=0.5) +
  ggtitle("Home Electricity in 15-Minute Intervals") +
  ylab("Electricity Consumption (kWh)")

```

Example day with EV charging

```{r}
sample_day <- hourly_electricity %>% 
  filter(solar_year == 1 | solar_year == 2) %>% 
  filter(yday == 161) %>% 
  group_by(date, solar_year, yday, time) %>% 
  summarize(produced_kWh = mean(produced_kWh), consumed_kWh = mean(consumed_kWh), net_kWh = mean(net_kWh)) %>% 
  arrange(date, solar_year, yday, time)

ggplot(sample_day, aes(time, consumed_kWh, group=solar_year, color=solar_year)) +
  #facet_grid(rows = vars(month)) +
  geom_point(size=0.5) +
  ggtitle("Home Electricity in 15-Minute Intervals") +
  ylab("Electricity Consumption (kWh)")

```

Ok, let's detect and remove EV charging sessions. 

Bayesian change point example
```{r}
# library(onlineBcp)
x <- c(rnorm(10, 2.6, 0.2), rnorm(1, 2.6/2, 0.2) + rnorm(1, .2, .15), rnorm(86, .2, .15))
bcp <- online_cp(x)
summary(bcp)

bcp

```

Test on electricity consumption data
```{r}
# library(onlineBcp)

#Select a sample to test
sample <- hourly_electricity %>% 
  mutate(solar_day = interval(as_date(as.POSIXct('04/15/2022',format="%m/%d/%Y",tz=Sys.timezone())),as_date(date)) / days(1) ) %>%
  filter(solar_year == 2) %>% 
  filter(yday < 40 & yday > 35)

ggplot(sample, aes(datetime, consumed_kWh, group=solar_year, color=solar_year)) +
  #facet_grid(rows = vars(month)) +
  geom_point(size=0.5) +
  ggtitle("Home Electricity in 15-Minute Intervals") +
  ylab("Electricity Consumption (kWh)")
```


```{r}
min_time <- pull(sample %>% summarize(min_time = min(datetime)))

sample <- sample %>% 
  mutate(time_x = interval(min_time,datetime)/minutes(15)+1)

#Filter data to feed to model
x <- sample %>%  
  ungroup() %>% 
  select(consumed_kWh)

x <- x$consumed_kWh

#Run online Bayesian Change Point model 
#"Online" means in progress, not retroactive
bcp <- online_cp(x)

summary(bcp)

plot(summary(bcp))

```
Not too bad! It picks up the EV charging sessions, but it also picks up some other, presumably HVAC-related consumption changes. It looks like I should be able to weed these out easily. 


```{r}

#Select a sample to test
sample <- hourly_electricity %>% 
  mutate(solar_day = interval(as_date(as.POSIXct('04/15/2022',format="%m/%d/%Y",tz=Sys.timezone())),as_date(date)) / days(1) ) %>%
  filter(solar_year == 2)

ggplot(sample, aes(datetime, consumed_kWh, group=solar_year, color=solar_year)) +
  #facet_grid(rows = vars(month)) +
  geom_point(size=0.5) +
  ggtitle("Home Electricity in 15-Minute Intervals") +
  ylab("Electricity Consumption (kWh)")
```

```{r}

min_time <- pull(sample %>% summarize(min_time = min(datetime)))

sample <- sample %>% 
  mutate(time_x = interval(min_time,datetime)/minutes(15)+1)

#Filter data to feed to model
x <- sample %>%  
  ungroup() %>% 
  select(consumed_kWh)

x <- x$consumed_kWh

#Run online Bayesian Change Point model 
#"Online" means in progress, not retroactive
bcp <- online_cp(x)

summary(bcp)

plot(summary(bcp))
```


onlineBcp model seems to change efficacy when looking across different lengths of time. It works well on single days, but when looking at a week, it picks up noise from HVAC/other appliances, though still picking up the EV charging pretty well. When looking at longer periods, however, the EV charging intervals get lost in the noise.

Also, the original bcp package for R was deprecated, leaving the best option as onlineBCP. 
Maybe build a training data set with manually labeled EV charging intervals, then run a machine learning model on the rest of the data set? Use time series inputs. 

Identify streaks of high electricity usage (2 kWh in 15 minute interval and 2 kWh in previous and/or next interval)
```{r}
sample <- sample %>% 
  mutate(next_net = lead(consumed_kWh, order_by = datetime), prev_net = lag(consumed_kWh, order_by = datetime)) %>% 
  #mutate(next2 = lead(consumed_kWh, order_by = datetime, n = 2L), next3 = lead(consumed_kWh, order_by = datetime, n = 3L)) %>% 
  #mutate(nexthour_net = consumed_kWh + next_net + next2 + next3) %>% 
  mutate(streak = case_when(consumed_kWh > 2 & (next_net > 2 | prev_net > 2) ~ 1)) %>% 
  #mutate(streak_net = ) %>% 
  #mutate(start_of_ev_charge = case_when(streak == 1 & is.na(lag(streak, order_by = datetime)) ~ datetime), #& nexthour_net >= 10000
  #       end_of_ev_charge = case_when(streak == 1 & is.na(lead(streak, order_by = datetime)) ~ end), 
  #       ev_charge_time = 0, ev_charge_net = 0, ev_charge_min = 0, ev_charge_max = 0, customer_id = customer_id) %>% 
  arrange(datetime)

sample
```

Now, look at length of streaks to identify charging sessions. (not finished)

```{r eval=FALSE, include=FALSE}

sample %>% 
  mutate(start_of_streak = case_when(streak == 1 & lag()))


```


This doesn't work yet: 
```{r eval=FALSE, include=FALSE}
sample <- as.data.frame(sample)

i <- 2
while (i <= nrow(sample)) {
  sample[i]$start_of_ev_charge <- case_when(sample[i]$streak == 1 & is.na(sample[i-1]$streak) ~ sample[i]$start_of_ev_charge, 
                                       sample[i]$streak == 1 & !is.na(sample[i-1]$streak) ~ sample[i-1]$start_of_ev_charge)
  sample[i]$ev_charge_time <- case_when(sample[i]$streak == 1 & is.na(sample[i-1]$streak) ~ 15, 
                                       sample[i]$streak == 1 & !is.na(sample[i-1]$streak) ~ 15 + sample[i-1]$ev_charge_time)
  sample[i]$ev_charge_net <- case_when(sample[i]$streak == 1 & is.na(sample[i-1]$streak) ~ sample[i]$net_electricity_consumption, 
                                       sample[i]$streak == 1 & !is.na(sample[i-1]$streak) ~ sample[i]$net_electricity_consumption + sample[i-1]$ev_charge_net)
  
  sample[i]$ev_charge_min <- case_when(sample[i]$streak == 1 & is.na(sample[i-1]$streak) ~ sample[i]$net_electricity_consumption, 
                                       sample[i]$streak == 1 & !is.na(sample[i-1]$streak) & sample[i]$net_electricity_consumption <= sample[i-1]$ev_charge_min ~ sample[i]$net_electricity_consumption, 
                                       sample[i]$streak == 1 & !is.na(sample[i-1]$streak) & sample[i]$net_electricity_consumption > sample[i-1]$ev_charge_min ~ sample[i-1]$ev_charge_min )
  
  sample[i]$ev_charge_max <- case_when(sample[i]$streak == 1 & is.na(sample[i-1]$streak) ~ sample[i]$net_electricity_consumption, 
                                       sample[i]$streak == 1 & !is.na(sample[i-1]$streak) & sample[i]$net_electricity_consumption >= sample[i-1]$ev_charge_max ~ sample[i]$net_electricity_consumption, 
                                       sample[i]$streak == 1 & !is.na(sample[i-1]$streak) & sample[i]$net_electricity_consumption < sample[i-1]$ev_charge_max ~ sample[i-1]$ev_charge_max )
  
  i <- i + 1
}

#sample %>% 
#  filter(streak == 1)

#streaks<- sample %>% 
streaks_ind <- sample %>% 
  filter(!is.na(end_of_ev_charge)) %>% 
  select(customer_id, start_of_ev_charge, end_of_ev_charge, ev_charge_min, ev_charge_max)

streaks_ind


```


