---
title: "Time of Use Electricity Rate Analysis"
output: html_notebook
---

```{r message=FALSE, warning=FALSE}
#library(rmarkdown)
library(tidyverse)
library(lubridate)
library(hms)
```
Import and examine PPL 15 minute net meter interval data

```{r}
hourly_ppl_pivot <- read_csv("ppl_15mins.csv")

(hourly_ppl_net <- hourly_ppl_pivot %>% 
  filter(`Read Type` == "kWh Net") %>% 
  arrange(desc(date)))
```

Load Enphase data
```{r}
full_hist <- read_csv("enphase_production.csv")

(hourly_production <- full_hist %>% 
  rename(datetime = `Date/Time`, energy_produced_Wh = `Energy Produced (Wh)`) %>% 
  mutate(datetime = as.POSIXct(datetime,format="%m/%d/%Y %H:%M",tz=Sys.timezone())) %>% 
  mutate(date = date(datetime), time = as.hms(format(datetime, format = "%H:%M:%S")), month = month(datetime, label=TRUE), year = year(datetime), day = day(datetime), yday = yday(datetime), monthday = format(datetime, "%m-%d"), wday = wday(datetime, label=TRUE), equinox_day = (yday + 10) %% 365, equinox_group = floor((equinox_day+15)/30)*30) %>% 
    arrange(desc(datetime))
)

```


Net + Produced = Consumed
Start and End Dates of electricity data
```{r}
(hourly_electricity <- hourly_ppl_net %>% 
    inner_join(hourly_production, by = join_by(date,time))  %>% 
    mutate(consumed_kWh = kWh + energy_produced_Wh/1000, produced_kWh = energy_produced_Wh/1000)  %>% 
    rename(net_kWh = kWh) %>% 
    select(datetime, date, time, net_kWh, produced_kWh, consumed_kWh) %>%
    arrange(desc(datetime)))

hourly_electricity %>% 
  summarize(min(date), max(date))

```

Cost of fixed rate electricity over the full period. 
```{r}
fixed_rate <- 0.1004 + 0.04727 #PA average electricity rate, supply + delivery charges combined

hourly_electricity %>% 
  mutate(net_fixed = net_kWh * fixed_rate, consumed_fixed = consumed_kWh * fixed_rate) %>% 
  #group_by(year(date), month(date)) %>% 
  summarize(produced_kWh = sum(produced_kWh), consumed_kWh = sum(consumed_kWh), net_fixed_cost = sum(net_fixed), consumed_fixed_cost = sum(consumed_fixed))

```
Peak vs. Off-peak consumption/production

```{r}
electricity_by_time_jun_nov <- hourly_electricity %>% 
  filter(month(date) %in% c(6,7,8,9,10,11)) %>% 
  group_by(time) %>% 
  summarize(produced_kWh = sum(produced_kWh), consumed_kWh = sum(consumed_kWh), net_kWh = sum(consumed_kWh) - sum(produced_kWh)) %>% 
  mutate(produced_pct = produced_kWh/(consumed_kWh+produced_kWh))

ggplot(electricity_by_time_jun_nov, aes(x = time, y = produced_pct)) +
  geom_bar(stat="identity") +
  geom_vline(xintercept = as_hms("14:00:00"), colour="green", linetype = "longdash") + 
  geom_vline(xintercept = as_hms("18:00:00"), colour="red", linetype = "longdash") +
  ggtitle("Summer Production Pct")

electricity_by_time_dec_may <- hourly_electricity %>% 
  filter(month(date) %in% c(12,1,2,3,4,5)) %>% 
  group_by(time) %>% 
  summarize(produced_kWh = sum(produced_kWh), consumed_kWh = sum(consumed_kWh), net_kWh = sum(consumed_kWh) - sum(produced_kWh)) %>% 
  mutate(produced_pct = produced_kWh/(consumed_kWh+produced_kWh))

ggplot(electricity_by_time_dec_may, aes(x = time, y = produced_pct)) +
  geom_bar(stat="identity") +
  geom_vline(xintercept = as_hms("16:00:00"), colour="green", linetype = "longdash") + 
  geom_vline(xintercept = as_hms("20:00:00"), colour="red", linetype = "longdash") +
  ggtitle("Winter Production Pct")
```


```{r}
#monthly totals by peak/off-peak
peak_off_peak_electricity_by_month <- hourly_electricity %>% 
  mutate(year = year(date), month = month(date), year_month = format(date,"%Y-%m"),
    category = case_when(wday(datetime) %in% c(1,7) ~ "off_peak", #weekends
                                   hour(datetime) >= 14 & hour(datetime) < 18 & month(datetime) %in% c(6,7,8,9,10,11) ~ "peak", #JuneToNovPeak
                                   hour(datetime) >= 16 & hour(datetime) < 20 & month(datetime) %in% c(12,1,2,3,4,5) ~ "peak", #DecToMayPeak
                                   .default = "off_peak"
                                   ) 
         ) %>% 
  group_by(year, month, year_month, category) %>% 
  summarize(produced_kWh = sum(produced_kWh), consumed_kWh = sum(consumed_kWh), net_kWh = sum(consumed_kWh) - sum(produced_kWh)) %>% 
  arrange(year, month)

peak_off_peak_electricity_by_month


#peak/offpeak totals
peak_off_peak_electricity <- peak_off_peak_electricity_by_month %>% 
  group_by(category) %>% 
  summarize(produced_kWh = sum(produced_kWh), consumed_kWh = sum(consumed_kWh), net_kWh = sum(consumed_kWh) - sum(produced_kWh)) 

peak_off_peak_electricity %>% 
  summarize(produced_off_peak = sum(case_when(category=="off_peak" ~ produced_kWh, .default = 0)), 
         produced_peak = sum(case_when(category=="peak" ~ produced_kWh, .default = 0)), 
         consumed_off_peak = sum(case_when(category=="off_peak" ~ consumed_kWh, .default = 0)), 
         consumed_peak = sum(case_when(category=="peak" ~ consumed_kWh, .default = 0)), 
         net_off_peak = sum(case_when(category=="off_peak" ~ net_kWh, .default = 0)), 
         net_peak = sum(case_when(category=="peak" ~ net_kWh, .default = 0)), 
         ) %>% 
  mutate(produced_peak_pct = produced_peak/(produced_peak+produced_off_peak), 
         consumed_peak_pct = consumed_peak/(consumed_peak+consumed_off_peak))

#by month pivot
peak_off_peak_electricity_by_month_pivot <- peak_off_peak_electricity_by_month %>% 
  group_by(year, month, year_month) %>% 
  summarize(produced_off_peak = sum(case_when(category=="off_peak" ~ produced_kWh, .default = 0)), 
         produced_peak = sum(case_when(category=="peak" ~ produced_kWh, .default = 0)), 
         consumed_off_peak = sum(case_when(category=="off_peak" ~ consumed_kWh, .default = 0)), 
         consumed_peak = sum(case_when(category=="peak" ~ consumed_kWh, .default = 0)), 
         net_off_peak = sum(case_when(category=="off_peak" ~ net_kWh, .default = 0)), 
         net_peak = sum(case_when(category=="peak" ~ net_kWh, .default = 0)), 
         ) %>% 
  mutate(produced_peak_pct = produced_peak/(produced_peak+produced_off_peak), 
         consumed_peak_pct = consumed_peak/(consumed_peak+consumed_off_peak))

peak_off_peak_electricity_by_month_pivot

#Plot peak usage by month
ggplot(peak_off_peak_electricity_by_month, aes(x = year_month, y = consumed_kWh, fill = category)) +
  geom_bar(stat="identity") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggplot(peak_off_peak_electricity_by_month, aes(x = year_month, y = produced_kWh, fill = category)) +
  geom_bar(stat="identity") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggplot(peak_off_peak_electricity_by_month, aes(x = year_month, y = net_kWh, fill = category)) +
  geom_bar(stat="identity") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))


```

Time of Use rate cost over the full period

Step 1: Usage totaled
We total your usage and generation by each season. If you generated more than you used, we’ll add these to your bank. You’ll still receive a bill for basic service. If you used more than you generated, we move to Step 2.
- Might have to go back and group by season as well as month/year

Step 2: Current bill
In the same bill cycle, you may have usage to offset in one season or time period, and excess generation in another. If this is the case, we use your excess generation first to offset any usage you would be billed for.

Step 3: Bank applied
If you still have usage you would be billed for, we pull any existing or previously banked hours to offset your bill. Banked hours are used chronologically. You will be billed for any kilowatt-hours used above your banked amount.



How is my bank used?
If you generate more than you use, you’ll bank kilowatt-hours at seasonal on- and off-peak rates. Then if there is a month when you use more than you generate, we’ll use some of those banked hours to offset your energy costs.

Here is how we use your bank to offset your usage:

Banked hours are used chronologically. For example, June’s off-peak hours would be used first, then June’s on-peak hours.
Your usage is also offset chronologically. We offset your oldest on-peak hours first. Then your oldest off-peak hours.


```{r}
#https://www.pplelectric.com/my-account/shop-for-electricity/time-of-use-plan/tou-enrollment
peak_rate <- 0.12936 + 0.04727
off_peak_rate <- 0.09319 + 0.04727

#write_csv(peak_off_peak_electricity_by_month_pivot, "peak_off_peak_electricity_by_month_pivot.csv")

df <- peak_off_peak_electricity_by_month_pivot %>% 
  rowid_to_column(var='row_id') %>% 
  mutate(cum_off_peak_bank = 0, cum_peak_bank = 0) %>% 
  ungroup() %>% 
  slice_min(row_id, n=15)

#set variables net, net_peak, net_off_peak, peak_bank, off_peak_bank
net <- 0 
net_peak <- 0
net_off_peak <- 0
peak_bank <- 0 
off_peak_bank <- 0

i = 1
  
while(i <= nrow(df)) {
  
  #print(i)
  
  #pull net kWh values from this month
  net_peak <- pull(df %>% filter(row_id == i) %>% select(net_peak))
  net_off_peak <- pull(df %>% filter(row_id == i) %>% select(net_off_peak))
  net <- net_peak + net_off_peak

  if (net <= 0) {
    if (net_peak <= 0 & net_off_peak <= 0) {
      peak_bank = peak_bank + net_peak
      off_peak_bank = off_peak_bank + net_off_peak
    } 
    else if (net_peak <= 0 & net_off_peak > 0) {
      peak_bank = peak_bank + net
    }
    else if (net_off_peak <= 0 & net_peak > 0) {
      off_peak_bank = off_peak_bank + net
    }
  } else if (net > 0) {
    if (net_peak > 0 & net_off_peak > 0) { 
      peak_bank = peak_bank + net_peak
      off_peak_bank = off_peak_bank + net_off_peak
    } else if (net_peak <= 0 & net_off_peak > 0) {
      off_peak_bank = off_peak_bank + net
    } else if (net_off_peak <= 0 & net_peak > 0) { 
      peak_bank = peak_bank + net
    }
  } else if (off_peak_bank > 0 & peak_bank > 0) {
    #charge customer for each and reset to 0
    #fix these
    #df <- df %>% filter(row_id == i) %>% mutate(peak_charge = peak_rate*peak_bank)
    #df <- df %>% filter(row_id == i) %>% mutate(off_peak_charge = off_peak_rate*off_peak_bank)
    peak_bank <- 0 
    off_peak_bank <- 0
  } else if (peak_bank > 0) {
    #charge the customer and reset to 0
    df <- df %>% filter(row_id == i) %>% mutate(peak_charge = peak_rate*peak_bank)
    peak_bank <- 0 
    off_peak_bank <- 0
  }
  
  df <- rows_update(df, tibble(row_id = i, cum_off_peak_bank = off_peak_bank, cum_peak_bank = peak_bank), by = "row_id")
  #df$peak_bank <- peak_bank
  #df$off_peak_bank <- off_peak_bank
  
  #split by season (summer/winter)
  
  #if month == 5, then charge the customer and reset to 0 
  #peak_bank <- 0 
  #off_peak_bank <- 0
  
  i = i + 1
}

df

```


```{r}

```


Details to come back to:
```{r}
#add holidays to off_peak
  # New Year's Day
  # Presidents Day
  # Good Friday
  # Memorial Day
  # Independence Day
  # Labor Day
  # Thanksgiving Day
  # Day after Thanksgiving
  # Christmas Eve
  # Christmas Day

#separate seasons (even within a bill)

#account for rate changes

```








If TOU rate applied hourly (N/A): 
```{r eval=FALSE, include=FALSE}
hourly_electricity %>% 
  mutate(tou_rate = case_when(wday(datetime) %in% c(1,7) ~ off_peak_rate, #weekends
                                   hour(datetime) >= 14 & hour(datetime) < 18 & month(datetime) %in% c(6,7,8,9,10,11) ~ peak_rate, #JuneToNovPeak
                                   hour(datetime) >= 16 & hour(datetime) < 20 & month(datetime) %in% c(12,1,2,3,4,5) ~ peak_rate, #DecToMayPeak
                                   .default = off_peak_rate
                                   ) 
         ) %>% 
  #group_by(year(date), month(date)) %>%
  summarize(net_tou_cost = sum(net_kWh*tou_rate), consumed_tou_cost = sum(consumed_kWh*tou_rate))
  

```
Somewhat better to shift to TOU in both cases (with or without solar).


What if we shifted the EV charging load? 
- Detect EV charging load
- Give it the lower rate

Load Emporia charging 15 minute interval data
```{r eval=FALSE, include=FALSE}

min15_2023 <- read_csv("FA1BE4-Emporia_Garage-15MIN_20230609_20231231.csv", show_col_types = FALSE) %>%
                mutate(`Date/Time` = as.POSIXct(`Time Bucket (America/New_York)`,format="%m/%d/%Y %H:%M",tz=Sys.timezone()))
min15_2024 <- read_csv("FA1BE4-Emporia_Garage-15MIN_20240101_20240613.csv", show_col_types = FALSE) %>%
                mutate(`Date/Time` = as.POSIXct(`Time Bucket (America/New_York)`,format="%m/%d/%Y %H:%M",tz=Sys.timezone())) %>%
                filter(`Date/Time` < as.POSIXct('06/11/2024 00:00',format="%m/%d/%Y %H:%M",tz=Sys.timezone()) )
min15_2024_2 <- read_csv("FA1BE4-Emporia_Garage-15MIN_20240611_20240819.csv", show_col_types = FALSE) %>%
                mutate(`Date/Time` = as.POSIXct(`Time Bucket (America/New_York)`,format="%m/%d/%Y %H:%M",tz=Sys.timezone())) %>%
                filter(`Date/Time` < as.POSIXct('08/20/2024 00:00',format="%m/%d/%Y %H:%M",tz=Sys.timezone()) )

# min15_2023 <- read_csv("FA1BE4-Emporia_Garage-15MIN_20230609_20231231.csv", show_col_types = FALSE)
# min15_2024 <- read_csv("FA1BE4-Emporia_Garage-15MIN_20240101_20240613.csv", show_col_types = FALSE)
# min15_2024_2 <- read_csv("FA1BE4-Emporia_Garage-15MIN_20240611_20240819.csv", show_col_types = FALSE)

min15_charging <- rbind(min15_2023,min15_2024)
min15_charging <- rbind(min15_charging,min15_2024_2)

#min15_charging <- rbind(min15_2024,list(min15_2024,min15_2024_2))

#change data types
min15_charging <- min15_charging %>% 
  rename(datetime = `Time Bucket (America/New_York)`, kW = `Emporia Garage-EV (kWatts)`) %>% 
  #mutate(datetime = as_datetime(datetime, tz = 'America/New_York'))
  #mutate(datetime = as.POSIXct(datetime,format="%m/%d/%Y %H:%M:%S",tz='America/New_York'))
  mutate(datetime = mdy_hms(datetime, tz = 'America/New_York')) %>% 
  mutate(solar_year = factor(case_when(
                        datetime < as.POSIXct('04/16/2022 00:00',format="%m/%d/%Y %H:%M",tz=Sys.timezone()) ~ 0,
                        datetime >= as.POSIXct('04/16/2022 00:00',format="%m/%d/%Y %H:%M",tz=Sys.timezone())  &
                          datetime < as.POSIXct('04/16/2023 00:00',format="%m/%d/%Y %H:%M",tz=Sys.timezone()) ~ 1,
                        datetime >= as.POSIXct('04/16/2023 00:00',format="%m/%d/%Y %H:%M",tz=Sys.timezone())  &
                          datetime < as.POSIXct('04/16/2024 00:00',format="%m/%d/%Y %H:%M",tz=Sys.timezone()) ~ 2,
                        TRUE ~ 3)
         )) %>% 
  arrange(desc(datetime))

min15_charging <- min15_charging %>% 
  mutate(date = date(datetime), time = as_hms(datetime)) %>% 
  mutate(charging_kWh = kW/4) %>% 
  select (-c(solar_year,`Date/Time`,kW))

min15_charging
```

Check date range for available hourly charging data
```{r eval=FALSE, include=FALSE}
min15_charging %>% 
  summarize(min(datetime), max(datetime))

min_datetime <- pull(min15_charging %>% summarize(min = min(datetime)))
max_datetime <- pull(min15_charging %>% summarize(max = max(datetime)))
```

Run fixed and variable rates on those dates
Fixed rate cost on this date range.
```{r eval=FALSE, include=FALSE}
#Fixed rate

hourly_electricity %>% 
  filter(between(date, min_datetime, max_datetime)) %>% 
  arrange(desc(datetime))

hourly_electricity %>% 
  mutate(net_fixed_cost = net_kWh * fixed_rate, consumed_fixed_cost = consumed_kWh * fixed_rate) %>% 
  filter(between(datetime, min_datetime, max_datetime)) %>% 
  summarize(net_fixed_cost = sum(net_fixed_cost), consumed_fixed_cost = sum(consumed_fixed_cost))

```

TOU cost for this date range
```{r eval=FALSE, include=FALSE}
#TOU rate
hourly_electricity %>% 
  mutate(tou_rate = case_when(wday(datetime) %in% c(1,7) ~ off_peak_rate, #weekends
                                   hour(datetime) >= 14 & hour(datetime) < 18 & month(datetime) %in% c(6,7,8,9,10,11) ~ peak_rate, #JuneToNovPeak
                                   hour(datetime) >= 16 & hour(datetime) < 20 & month(datetime) %in% c(12,1,2,3,4,5) ~ peak_rate, #DecToMayPeak
                                   .default = off_peak_rate
                                   ) 
         ) %>% 
  filter(between(datetime, min_datetime, max_datetime)) %>% 
  summarize(net_tou_cost = sum(net_kWh*tou_rate), consumed_variable = sum(consumed_kWh*tou_rate))
  
```

Shift EV charging to lower rate
```{r eval=FALSE, include=FALSE}

#Join hourly_electricity to min15_charging
min15_electricity_charging <- hourly_electricity %>% 
  inner_join(min15_charging, by = join_by(date,time)) %>% 
  mutate(net_kWh_no_charging = net_kWh - charging_kWh, consumed_kWh_no_charging = consumed_kWh - charging_kWh) %>% 
  rename(datetime = datetime.x)

#fix github push issue

#variable rate
#shift charging load
#for peak hours, subtract ev charging kWh and add back at off-peak rate

min15_electricity_charging %>% 
  mutate(tou_rate = case_when(wday(datetime) %in% c(1,7) ~ off_peak_rate, #weekends
                                   hour(datetime) >= 14 & hour(datetime) < 18 & month(datetime) %in% c(6,7,8,9,10,11) ~ peak_rate, #JuneToNovPeak
                                   hour(datetime) >= 16 & hour(datetime) < 20 & month(datetime) %in% c(12,1,2,3,4,5) ~ peak_rate, #DecToMayPeak
                                   .default = off_peak_rate
                                   ) 
         ) %>% 
  filter(between(datetime, min_datetime, max_datetime)) %>% 
  summarize(net_tou_cost = sum(net_kWh_no_charging*tou_rate + charging_kWh*off_peak_rate), consumed_tou_cost = sum(consumed_kWh_no_charging*tou_rate + charging_kWh*off_peak_rate))
  

```

Looks like it saves even more. Over this ~14 month period, we would have saved about $70 more with a TOU rate that and EV charging load shifting.

Plot of energy usage by hour
- Split by months
```{r eval=FALSE, include=FALSE}

by_month <- hourly_electricity %>% 
  filter(between(datetime, min_datetime, max_datetime)) %>% 
  mutate(month = as.factor(month(date))) %>% 
  group_by(month, time) %>% 
  summarize(n = n(), net_kWh = mean(net_kWh), produced_kWh = mean(produced_kWh), consumed_kWh = mean(consumed_kWh))

by_month %>% 
  group_by(month) %>% 
  summarize(net_kWh = sum(net_kWh), produced_kWh = sum(produced_kWh), consumed_kWh = sum(consumed_kWh)) %>% 
  arrange(desc(net_kWh))

ggplot(by_month, aes(time, net_kWh, color=month)) +
  geom_point() 

ggplot(by_month, aes(time, consumed_kWh, color=month)) +
  geom_point()

ggplot(by_month, aes(time, produced_kWh, color=month)) +
  geom_point()

jun_nov_electricity <- hourly_electricity %>% 
  filter(between(datetime, min_datetime, max_datetime), month(date) %in% c(6,7,8,9,10,11)) %>% 
  group_by(time) %>% 
  summarize(n = n(), net_kWh = mean(net_kWh), produced_kWh = mean(produced_kWh), consumed_kWh = mean(consumed_kWh))

jun_nov_electricity <- hourly_electricity %>% 
  filter(between(datetime, min_datetime, max_datetime), month(date) %in% c(6,7,8,9,10,11)) %>% 
  group_by(time) %>% 
  summarize(n = n(), net_kWh = mean(net_kWh), produced_kWh = mean(produced_kWh), consumed_kWh = mean(consumed_kWh))

ggplot(jun_nov_electricity, aes(time, net_kWh)) +
  geom_point() + 
  geom_vline(xintercept = as_hms("14:00:00"), colour="green", linetype = "longdash") + 
  geom_vline(xintercept = as_hms("18:00:00"), colour="red", linetype = "longdash") +
  ggtitle("Jun to Nov Electricity Load")

ggplot(jun_nov_electricity, aes(time, consumed_kWh)) +
  geom_point() + 
  geom_vline(xintercept = as_hms("14:00:00"), colour="green", linetype = "longdash") + 
  geom_vline(xintercept = as_hms("18:00:00"), colour="red", linetype = "longdash") +
  ggtitle("Jun to Nov Electricity Load")

dec_may_electricity <- hourly_electricity %>% 
  filter(between(datetime, min_datetime, max_datetime), month(date) %in% c(12,1,2,3,4,5)) %>% 
  group_by(time) %>% 
  summarize(n = n(), net_kWh = mean(net_kWh), produced_kWh = mean(produced_kWh), consumed_kWh = mean(consumed_kWh))

ggplot(dec_may_electricity, aes(time, net_kWh)) +
  geom_point() + 
  geom_vline(xintercept = as_hms("16:00:00"), colour="green", linetype = "longdash") + 
  geom_vline(xintercept = as_hms("20:00:00"), colour="red", linetype = "longdash") +
  ggtitle("Dec to May Electricity Load")

ggplot(dec_may_electricity, aes(time, consumed_kWh)) +
  geom_point() + 
  geom_vline(xintercept = as_hms("16:00:00"), colour="green", linetype = "longdash") + 
  geom_vline(xintercept = as_hms("20:00:00"), colour="red", linetype = "longdash") +
  ggtitle("Dec to May Electricity Load")


```

EV charging by hour


Load without charging 
```{r eval=FALSE, include=FALSE}

jun_nov_electricity_no_charging <- min15_electricity_charging %>% 
  filter(between(datetime, min_datetime, max_datetime), month(date) %in% c(6,7,8,9,10,11)) %>% 
  mutate(period = "Jun to Nov") %>% 
  group_by(period, time) %>% 
  summarize(n = n(), net_kWh = mean(net_kWh_no_charging), produced_kWh = mean(produced_kWh), consumed_kWh = mean(consumed_kWh_no_charging))

ggplot(jun_nov_electricity_no_charging, aes(time, net_kWh)) +
  geom_point() + 
  geom_vline(xintercept = as_hms("14:00:00"), colour="green", linetype = "longdash") + 
  geom_vline(xintercept = as_hms("18:00:00"), colour="red", linetype = "longdash") +
  ggtitle("Jun to Nov Electricity Load - No Charging")

ggplot(jun_nov_electricity_no_charging, aes(time, consumed_kWh)) +
  geom_point() + 
  geom_vline(xintercept = as_hms("14:00:00"), colour="green", linetype = "longdash") + 
  geom_vline(xintercept = as_hms("18:00:00"), colour="red", linetype = "longdash") +
  ggtitle("Jun to Nov Electricity Load - No Charging")

dec_may_electricity_no_charging <- min15_electricity_charging %>% 
  filter(between(datetime, min_datetime, max_datetime), month(date) %in% c(12,1,2,3,4,5)) %>% 
  mutate(period = "Dec to May") %>% 
  group_by(period, time) %>% 
  summarize(n = n(), net_kWh = mean(net_kWh_no_charging), produced_kWh = mean(produced_kWh), consumed_kWh = mean(consumed_kWh_no_charging))

ggplot(dec_may_electricity_no_charging, aes(time, net_kWh)) +
  geom_point() + 
  geom_vline(xintercept = as_hms("16:00:00"), colour="green", linetype = "longdash") + 
  geom_vline(xintercept = as_hms("20:00:00"), colour="red", linetype = "longdash") +
  ggtitle("Dec to May Electricity Load - No Charging")

ggplot(dec_may_electricity_no_charging, aes(time, consumed_kWh)) +
  geom_point() + 
  geom_vline(xintercept = as_hms("16:00:00"), colour="green", linetype = "longdash") + 
  geom_vline(xintercept = as_hms("20:00:00"), colour="red", linetype = "longdash") +
  ggtitle("Dec to May Electricity Load - No Charging")

electricity_no_charging_by_period <- rbind(jun_nov_electricity_no_charging, dec_may_electricity_no_charging)

ggplot(electricity_no_charging_by_period, aes(time, consumed_kWh, group=period, color=period)) +
  geom_point() + 
  geom_vline(xintercept = as_hms("16:00:00"), color="red", linetype = "longdash") + 
  geom_vline(xintercept = as_hms("20:00:00"), color="red", linetype = "longdash") +
  geom_vline(xintercept = as_hms("14:00:00"), color="light blue", linetype = "longdash") + 
  geom_vline(xintercept = as_hms("18:00:00"), color="light blue", linetype = "longdash")

```



```{r eval=FALSE, include=FALSE}
min15_electricity_charging 

jun_nov_electricity <- hourly_electricity %>% 
  filter(between(datetime, min_datetime, max_datetime), month(date) %in% c(6,7,8,9,10,11)) %>% 
  group_by(time) %>% 
  summarize(n = n(), net_kWh = mean(net_kWh), produced_kWh = mean(produced_kWh), consumed_kWh = mean(consumed_kWh))

ggplot(jun_nov_electricity, aes(time, net_kWh)) +
  geom_point() + 
  geom_vline(xintercept = as_hms("14:00:00"), colour="green", linetype = "longdash") + 
  geom_vline(xintercept = as_hms("18:00:00"), colour="red", linetype = "longdash")

ggplot(jun_nov_electricity, aes(time, consumed_kWh)) +
  geom_point() + 
  geom_vline(xintercept = as_hms("14:00:00"), colour="green", linetype = "longdash") + 
  geom_vline(xintercept = as_hms("18:00:00"), colour="red", linetype = "longdash")

dec_may_electricity <- hourly_electricity %>% 
  filter(between(datetime, min_datetime, max_datetime), month(date) %in% c(12,1,2,3,4,5)) %>% 
  group_by(time) %>% 
  summarize(n = n(), net_kWh = mean(net_kWh), produced_kWh = mean(produced_kWh), consumed_kWh = mean(consumed_kWh))

ggplot(dec_may_electricity, aes(time, net_kWh)) +
  geom_point() + 
  geom_vline(xintercept = as_hms("16:00:00"), colour="green", linetype = "longdash") + 
  geom_vline(xintercept = as_hms("20:00:00"), colour="red", linetype = "longdash")

ggplot(dec_may_electricity, aes(time, consumed_kWh)) +
  geom_point() + 
  geom_vline(xintercept = as_hms("16:00:00"), colour="green", linetype = "longdash") + 
  geom_vline(xintercept = as_hms("20:00:00"), colour="red", linetype = "longdash")

```



Plot of peak/off peak energy usage (stacked bar chart?)
- Same plot with shifted load
- Split by months

On peak vs off peak kWh by day/month


Charging by time of day? Winter vs. summer? 
```{r eval=FALSE, include=FALSE}
seasonal_charging_times <- min15_electricity_charging %>% 
  mutate(season = case_when(month(date) %in% c(6,7,8,9,10,11) ~ 'JunNov', 
                            .default = 'DecMay')) %>% 
  group_by(season, time) %>% 
  summarize(charging_kWh = mean(charging_kWh))

ggplot(seasonal_charging_times, aes(time, charging_kWh, group=season, color=season)) +
  geom_point() + 
  geom_vline(xintercept = as_hms("16:00:00"), color="red", linetype = "longdash") + 
  geom_vline(xintercept = as_hms("20:00:00"), color="red", linetype = "longdash") +
  geom_vline(xintercept = as_hms("14:00:00"), color="light blue", linetype = "longdash") + 
  geom_vline(xintercept = as_hms("18:00:00"), color="light blue", linetype = "longdash")
```


Plot of my average energy usage in "summer" vs. "winter"
```{r eval=FALSE, include=FALSE}
seasonal_electricity <- hourly_electricity %>% 
  filter(between(datetime, min_datetime, max_datetime)) %>% 
  mutate(season = case_when(month(date) %in% c(6,7,8,9,10,11) ~ 'JunNov', 
                            .default = 'DecMay')) %>% 
  group_by(season, time) %>% 
  summarize(n = n(), net_kWh = mean(net_kWh), produced_kWh = mean(produced_kWh), consumed_kWh = mean(consumed_kWh))

ggplot(seasonal_electricity, aes(time, net_kWh, group=season, color=season)) +
  geom_point() + 
  geom_vline(xintercept = as_hms("16:00:00"), color="red", linetype = "longdash") + 
  geom_vline(xintercept = as_hms("20:00:00"), color="red", linetype = "longdash") +
  geom_vline(xintercept = as_hms("14:00:00"), color="light blue", linetype = "longdash") + 
  geom_vline(xintercept = as_hms("18:00:00"), color="light blue", linetype = "longdash")
  
```



plot of energy cost by day on each plan. 
  - Total by month

```{r eval=FALSE, include=FALSE}

#fixed rate
fixed_cost <- hourly_electricity %>% 
  mutate(rate_plan = 'fixed', net_fixed_cost = net_kWh * fixed_rate, consumed_fixed_cost = consumed_kWh * fixed_rate) %>% 
  filter(between(datetime, min_datetime, max_datetime)) %>% 
  group_by(rate_plan, date) %>% 
  summarize(net_cost = sum(net_fixed_cost), consumed_cost = sum(consumed_fixed_cost))


#TOU rate
tou_cost <- hourly_electricity %>% 
  mutate(rate_plan = 'tou', tou_rate = case_when(wday(datetime) %in% c(1,7) ~ off_peak_rate, #weekends
                                   hour(datetime) >= 14 & hour(datetime) < 18 & month(datetime) %in% c(6,7,8,9,10,11) ~ peak_rate, #JuneToNovPeak
                                   hour(datetime) >= 16 & hour(datetime) < 20 & month(datetime) %in% c(12,1,2,3,4,5) ~ peak_rate, #DecToMayPeak
                                   .default = off_peak_rate
                                   ) 
         ) %>% 
  filter(between(datetime, min_datetime, max_datetime)) %>% 
  group_by(rate_plan, date) %>% 
  summarize(net_cost = sum(net_kWh*tou_rate), consumed_cost = sum(consumed_kWh*tou_rate))
  

#TOU rate
#shift charging load
#for peak hours, subtract ev charging kWh and add back at off-peak rate

tou_adjusted_cost <- min15_electricity_charging %>% 
  mutate(rate_plan = 'tou_adj', tou_rate = case_when(wday(datetime) %in% c(1,7) ~ off_peak_rate, #weekends
                                   hour(datetime) >= 14 & hour(datetime) < 18 & month(datetime) %in% c(6,7,8,9,10,11) ~ peak_rate, #JuneToNovPeak
                                   hour(datetime) >= 16 & hour(datetime) < 20 & month(datetime) %in% c(12,1,2,3,4,5) ~ peak_rate, #DecToMayPeak
                                   .default = off_peak_rate
                                   ) 
         ) %>% 
  filter(between(datetime, min_datetime, max_datetime)) %>% 
  group_by(rate_plan, date) %>% 
  summarize(net_cost = sum(net_kWh_no_charging*tou_rate + charging_kWh*off_peak_rate), consumed_cost = sum(consumed_kWh_no_charging*tou_rate + charging_kWh*off_peak_rate))
  
cost_by_date <- rbind(fixed_cost,tou_cost,tou_adjusted_cost)

cost_by_month <- cost_by_date %>% 
  mutate(month = as_factor(month(date))) %>% 
  group_by(rate_plan, month) %>% 
  summarize(net_cost = mean(net_cost), consumed_cost = mean(consumed_cost))

ggplot(cost_by_month, aes(month, consumed_cost, group=rate_plan, color=rate_plan)) +
  geom_point()

cost_by_month
```


