---
title: "Variable Electricity Rate Analysis"
output: html_notebook
---

```{r}
#library(rmarkdown)
library(tidyverse)
library(lubridate)
library(hms)
```
Import and examine PPL 15 minute net meter interval data

```{r}
hourly_ppl_pivot <- read_csv("ppl_15mins.csv")

(hourly_ppl_net <- hourly_ppl_pivot %>% 
  filter(`Read Type` == "kWh Net") %>% 
  arrange(desc(date)))
```

Load Enphase data
```{r}
full_hist <- read_csv("enphase_production.csv")

(hourly_production <- full_hist %>% 
  rename(datetime = `Date/Time`, energy_produced_Wh = `Energy Produced (Wh)`) %>% 
  mutate(datetime = as.POSIXct(datetime,format="%m/%d/%Y %H:%M",tz=Sys.timezone())) %>% 
  mutate(date = date(datetime), time = as.hms(format(datetime, format = "%H:%M:%S")), month = month(datetime, label=TRUE), year = year(datetime), day = day(datetime), yday = yday(datetime), monthday = format(datetime, "%m-%d"), wday = wday(datetime, label=TRUE), equinox_day = (yday + 10) %% 365, equinox_group = floor((equinox_day+15)/30)*30) %>% 
    arrange(desc(datetime))
)

```


Net + Produced = Consumed
Start and End Dates of electricity data
```{r}
(hourly_electricity <- hourly_ppl_net %>% 
    inner_join(hourly_production, by = join_by(date,time))  %>% 
    mutate(consumed_kWh = kWh + energy_produced_Wh/1000, produced_kWh = energy_produced_Wh/1000)  %>% 
    rename(net_kWh = kWh) %>% 
    select(datetime, date, time, net_kWh, produced_kWh, consumed_kWh) %>%
    arrange(desc(datetime)))

hourly_electricity %>% 
  summarize(min(date), max(date))

```

Cost of fixed rate electricity over the full period. 
```{r}
fixed_rate <- 0.1004 + 0.04727 #PA average electricity rate, supply + delivery charges combined

hourly_electricity %>% 
  mutate(net_fixed = net_kWh * fixed_rate, consumed_fixed = consumed_kWh * fixed_rate) %>% 
  #group_by(year(date), month(date)) %>% 
  summarize(net_fixed = sum(net_fixed), consumed_fixed = sum(consumed_fixed))

```
Time of Use rate cost over the full period
```{r}
#https://www.pplelectric.com/my-account/shop-for-electricity/time-of-use-plan/tou-enrollment
peak_rate <- 0.12936 + 0.04727
off_peak_rate <- 0.09319 + 0.04727

#add holidays to off_peak
  # New Year's Day
  # Presidents Day
  # Good Friday
  # Memorial Day
  # Independence Day
  # Labor Day
  # Thanksgiving Day
  # Day after Thanksgiving
  # Christmas Eve
  # Christmas Day

hourly_electricity %>% 
  mutate(tou_rate = case_when(wday(datetime) %in% c(1,7) ~ off_peak_rate, #weekends
                                   hour(datetime) >= 14 & hour(datetime) < 18 & month(datetime) %in% c(6,7,8,9,10,11) ~ peak_rate, #JuneToNovPeak
                                   hour(datetime) >= 16 & hour(datetime) < 20 & month(datetime) %in% c(12,1,2,3,4,5) ~ peak_rate, #DecToMayPeak
                                   .default = off_peak_rate
                                   ) 
         ) %>% 
  #group_by(year(date), month(date)) %>%
  summarize(net_tou_cost = sum(net_kWh*tou_rate), consumed_tou_cost = sum(consumed_kWh*tou_rate))
  

```

Somewhat better to shift to TOU in both cases (with or without solar).


What if we shifted the EV charging load? 
- Detect EV charging load
- Give it the lower rate

Load Emporia charging 15 minute interval data
```{r}

min15_2023 <- read_csv("FA1BE4-Emporia_Garage-15MIN_20230609_20231231.csv", show_col_types = FALSE) %>%
                mutate(`Date/Time` = as.POSIXct(`Time Bucket (America/New_York)`,format="%m/%d/%Y %H:%M",tz=Sys.timezone()))
min15_2024 <- read_csv("FA1BE4-Emporia_Garage-15MIN_20240101_20240613.csv", show_col_types = FALSE) %>%
                mutate(`Date/Time` = as.POSIXct(`Time Bucket (America/New_York)`,format="%m/%d/%Y %H:%M",tz=Sys.timezone())) %>%
                filter(`Date/Time` < as.POSIXct('06/11/2024 00:00',format="%m/%d/%Y %H:%M",tz=Sys.timezone()) )
min15_2024_2 <- read_csv("FA1BE4-Emporia_Garage-15MIN_20240611_20240819.csv", show_col_types = FALSE) %>%
                mutate(`Date/Time` = as.POSIXct(`Time Bucket (America/New_York)`,format="%m/%d/%Y %H:%M",tz=Sys.timezone())) %>%
                filter(`Date/Time` < as.POSIXct('08/20/2024 00:00',format="%m/%d/%Y %H:%M",tz=Sys.timezone()) )

# min15_2023 <- read_csv("FA1BE4-Emporia_Garage-15MIN_20230609_20231231.csv", show_col_types = FALSE)
# min15_2024 <- read_csv("FA1BE4-Emporia_Garage-15MIN_20240101_20240613.csv", show_col_types = FALSE)
# min15_2024_2 <- read_csv("FA1BE4-Emporia_Garage-15MIN_20240611_20240819.csv", show_col_types = FALSE)

min15_charging <- rbind(min15_2023,min15_2024)
min15_charging <- rbind(min15_charging,min15_2024_2)

#min15_charging <- rbind(min15_2024,list(min15_2024,min15_2024_2))

#change data types
min15_charging <- min15_charging %>% 
  rename(datetime = `Time Bucket (America/New_York)`, kW = `Emporia Garage-EV (kWatts)`) %>% 
  #mutate(datetime = as_datetime(datetime, tz = 'America/New_York'))
  #mutate(datetime = as.POSIXct(datetime,format="%m/%d/%Y %H:%M:%S",tz='America/New_York'))
  mutate(datetime = mdy_hms(datetime, tz = 'America/New_York')) %>% 
  mutate(solar_year = factor(case_when(
                        datetime < as.POSIXct('04/16/2022 00:00',format="%m/%d/%Y %H:%M",tz=Sys.timezone()) ~ 0,
                        datetime >= as.POSIXct('04/16/2022 00:00',format="%m/%d/%Y %H:%M",tz=Sys.timezone())  &
                          datetime < as.POSIXct('04/16/2023 00:00',format="%m/%d/%Y %H:%M",tz=Sys.timezone()) ~ 1,
                        datetime >= as.POSIXct('04/16/2023 00:00',format="%m/%d/%Y %H:%M",tz=Sys.timezone())  &
                          datetime < as.POSIXct('04/16/2024 00:00',format="%m/%d/%Y %H:%M",tz=Sys.timezone()) ~ 2,
                        TRUE ~ 3)
         )) %>% 
  arrange(desc(datetime))

min15_charging <- min15_charging %>% 
  mutate(date = date(datetime), time = as_hms(datetime)) %>% 
  mutate(charging_kWh = kW/4) %>% 
  select (-c(solar_year,`Date/Time`,kW))

min15_charging
```

Check date range for available hourly charging data
```{r}
min15_charging %>% 
  summarize(min(datetime), max(datetime))

min_datetime <- pull(min15_charging %>% summarize(min = min(datetime)))
max_datetime <- pull(min15_charging %>% summarize(max = max(datetime)))
```

Run fixed and variable rates on those dates
Fixed rate cost on this data rate.
```{r}
#Fixed rate

hourly_electricity %>% 
  filter(between(date, min_datetime, max_datetime)) %>% 
  arrange(desc(datetime))

hourly_electricity %>% 
  mutate(net_fixed_cost = net_kWh * fixed_rate, consumed_fixed_cost = consumed_kWh * fixed_rate) %>% 
  filter(between(datetime, min_datetime, max_datetime)) %>% 
  summarize(net_fixed_cost = sum(net_fixed_cost), consumed_fixed_cost = sum(consumed_fixed_cost))

```

TOU cost for this date range
```{r}
#variable rate
hourly_electricity %>% 
  mutate(tou_rate = case_when(wday(datetime) %in% c(1,7) ~ off_peak_rate, #weekends
                                   hour(datetime) >= 14 & hour(datetime) < 18 & month(datetime) %in% c(6,7,8,9,10,11) ~ peak_rate, #JuneToNovPeak
                                   hour(datetime) >= 16 & hour(datetime) < 20 & month(datetime) %in% c(12,1,2,3,4,5) ~ peak_rate, #DecToMayPeak
                                   .default = off_peak_rate
                                   ) 
         ) %>% 
  filter(between(datetime, min_datetime, max_datetime)) %>% 
  summarize(net_tou_cost = sum(net_kWh*tou_rate), consumed_variable = sum(consumed_kWh*tou_rate))
  
```

Shift EV charging to lower rate
```{r}

#Join hourly_electricity to min15_charging
min15_electricity_charging <- hourly_electricity %>% 
  inner_join(min15_charging, by = join_by(date,time)) %>% 
  mutate(net_kWh_no_charging = net_kWh - charging_kWh, consumed_kWh_no_charging = consumed_kWh - charging_kWh) %>% 
  rename(datetime = datetime.x)

#fix github push issue

#variable rate
#shift charging load
#for peak hours, subtract ev charging kWh and add back at off-peak rate

min15_electricity_charging %>% 
  mutate(tou_rate = case_when(wday(datetime) %in% c(1,7) ~ off_peak_rate, #weekends
                                   hour(datetime) >= 14 & hour(datetime) < 18 & month(datetime) %in% c(6,7,8,9,10,11) ~ peak_rate, #JuneToNovPeak
                                   hour(datetime) >= 16 & hour(datetime) < 20 & month(datetime) %in% c(12,1,2,3,4,5) ~ peak_rate, #DecToMayPeak
                                   .default = off_peak_rate
                                   ) 
         ) %>% 
  filter(between(datetime, min_datetime, max_datetime)) %>% 
  summarize(net_tou_cost = sum(net_kWh_no_charging*tou_rate + charging_kWh*off_peak_rate), consumed_tou_cost = sum(consumed_kWh_no_charging*tou_rate + charging_kWh*off_peak_rate))
  

```

Looks like it saves even more. Over this ~14 month period, we would have saved about $70 more with a TOU rate that and EV charging load shifting.

Plot of energy usage by hour
- Split by months
```{r}

by_month <- hourly_electricity %>% 
  filter(between(datetime, min_datetime, max_datetime)) %>% 
  mutate(month = as.factor(month(date))) %>% 
  group_by(month, time) %>% 
  summarize(n = n(), net_kWh = mean(net_kWh), produced_kWh = mean(produced_kWh), consumed_kWh = mean(consumed_kWh))

ggplot(by_month, aes(time, net_kWh, color=month)) +
  geom_point() 

ggplot(by_month, aes(time, consumed_kWh, color=month)) +
  geom_point()

ggplot(by_month, aes(time, produced_kWh, color=month)) +
  geom_point()

jun_nov_electricity <- hourly_electricity %>% 
  filter(between(datetime, min_datetime, max_datetime), month(date) %in% c(6,7,8,9,10,11)) %>% 
  group_by(time) %>% 
  summarize(n = n(), net_kWh = mean(net_kWh), produced_kWh = mean(produced_kWh), consumed_kWh = mean(consumed_kWh))

jun_nov_electricity <- hourly_electricity %>% 
  filter(between(datetime, min_datetime, max_datetime), month(date) %in% c(6,7,8,9,10,11)) %>% 
  group_by(time) %>% 
  summarize(n = n(), net_kWh = mean(net_kWh), produced_kWh = mean(produced_kWh), consumed_kWh = mean(consumed_kWh))

ggplot(jun_nov_electricity, aes(time, net_kWh)) +
  geom_point() + 
  geom_vline(xintercept = as_hms("14:00:00"), colour="green", linetype = "longdash") + 
  geom_vline(xintercept = as_hms("18:00:00"), colour="red", linetype = "longdash") +
  ggtitle("Jun to Nov Electricity Load")

ggplot(jun_nov_electricity, aes(time, consumed_kWh)) +
  geom_point() + 
  geom_vline(xintercept = as_hms("14:00:00"), colour="green", linetype = "longdash") + 
  geom_vline(xintercept = as_hms("18:00:00"), colour="red", linetype = "longdash") +
  ggtitle("Jun to Nov Electricity Load")

dec_may_electricity <- hourly_electricity %>% 
  filter(between(datetime, min_datetime, max_datetime), month(date) %in% c(12,1,2,3,4,5)) %>% 
  group_by(time) %>% 
  summarize(n = n(), net_kWh = mean(net_kWh), produced_kWh = mean(produced_kWh), consumed_kWh = mean(consumed_kWh))

ggplot(dec_may_electricity, aes(time, net_kWh)) +
  geom_point() + 
  geom_vline(xintercept = as_hms("16:00:00"), colour="green", linetype = "longdash") + 
  geom_vline(xintercept = as_hms("20:00:00"), colour="red", linetype = "longdash") +
  ggtitle("Dec to May Electricity Load")

ggplot(dec_may_electricity, aes(time, consumed_kWh)) +
  geom_point() + 
  geom_vline(xintercept = as_hms("16:00:00"), colour="green", linetype = "longdash") + 
  geom_vline(xintercept = as_hms("20:00:00"), colour="red", linetype = "longdash") +
  ggtitle("Dec to May Electricity Load")


```

EV charging by hour


Load without charging 
```{r}

jun_nov_electricity_no_charging <- min15_electricity_charging %>% 
  filter(between(datetime, min_datetime, max_datetime), month(date) %in% c(6,7,8,9,10,11)) %>% 
  mutate(period = "Jun to Nov") %>% 
  group_by(period, time) %>% 
  summarize(n = n(), net_kWh = mean(net_kWh_no_charging), produced_kWh = mean(produced_kWh), consumed_kWh = mean(consumed_kWh_no_charging))

ggplot(jun_nov_electricity_no_charging, aes(time, net_kWh)) +
  geom_point() + 
  geom_vline(xintercept = as_hms("14:00:00"), colour="green", linetype = "longdash") + 
  geom_vline(xintercept = as_hms("18:00:00"), colour="red", linetype = "longdash") +
  ggtitle("Jun to Nov Electricity Load - No Charging")

ggplot(jun_nov_electricity_no_charging, aes(time, consumed_kWh)) +
  geom_point() + 
  geom_vline(xintercept = as_hms("14:00:00"), colour="green", linetype = "longdash") + 
  geom_vline(xintercept = as_hms("18:00:00"), colour="red", linetype = "longdash") +
  ggtitle("Jun to Nov Electricity Load - No Charging")

dec_may_electricity_no_charging <- min15_electricity_charging %>% 
  filter(between(datetime, min_datetime, max_datetime), month(date) %in% c(12,1,2,3,4,5)) %>% 
  mutate(period = "Dec to May") %>% 
  group_by(period, time) %>% 
  summarize(n = n(), net_kWh = mean(net_kWh_no_charging), produced_kWh = mean(produced_kWh), consumed_kWh = mean(consumed_kWh_no_charging))

ggplot(dec_may_electricity_no_charging, aes(time, net_kWh)) +
  geom_point() + 
  geom_vline(xintercept = as_hms("16:00:00"), colour="green", linetype = "longdash") + 
  geom_vline(xintercept = as_hms("20:00:00"), colour="red", linetype = "longdash") +
  ggtitle("Dec to May Electricity Load - No Charging")

ggplot(dec_may_electricity_no_charging, aes(time, consumed_kWh)) +
  geom_point() + 
  geom_vline(xintercept = as_hms("16:00:00"), colour="green", linetype = "longdash") + 
  geom_vline(xintercept = as_hms("20:00:00"), colour="red", linetype = "longdash") +
  ggtitle("Dec to May Electricity Load - No Charging")

electricity_no_charging_by_period <- rbind(jun_nov_electricity_no_charging, dec_may_electricity_no_charging)

ggplot(electricity_no_charging_by_period, aes(time, consumed_kWh, group=period, color=period)) +
  geom_point()


```



```{r}
min15_electricity_charging 

jun_nov_electricity <- hourly_electricity %>% 
  filter(between(datetime, min_datetime, max_datetime), month(date) %in% c(6,7,8,9,10,11)) %>% 
  group_by(time) %>% 
  summarize(n = n(), net_kWh = mean(net_kWh), produced_kWh = mean(produced_kWh), consumed_kWh = mean(consumed_kWh))

ggplot(jun_nov_electricity, aes(time, net_kWh)) +
  geom_point() + 
  geom_vline(xintercept = as_hms("14:00:00"), colour="green", linetype = "longdash") + 
  geom_vline(xintercept = as_hms("18:00:00"), colour="red", linetype = "longdash")

ggplot(jun_nov_electricity, aes(time, consumed_kWh)) +
  geom_point() + 
  geom_vline(xintercept = as_hms("14:00:00"), colour="green", linetype = "longdash") + 
  geom_vline(xintercept = as_hms("18:00:00"), colour="red", linetype = "longdash")

dec_may_electricity <- hourly_electricity %>% 
  filter(between(datetime, min_datetime, max_datetime), month(date) %in% c(12,1,2,3,4,5)) %>% 
  group_by(time) %>% 
  summarize(n = n(), net_kWh = mean(net_kWh), produced_kWh = mean(produced_kWh), consumed_kWh = mean(consumed_kWh))

ggplot(dec_may_electricity, aes(time, net_kWh)) +
  geom_point() + 
  geom_vline(xintercept = as_hms("16:00:00"), colour="green", linetype = "longdash") + 
  geom_vline(xintercept = as_hms("20:00:00"), colour="red", linetype = "longdash")

ggplot(dec_may_electricity, aes(time, consumed_kWh)) +
  geom_point() + 
  geom_vline(xintercept = as_hms("16:00:00"), colour="green", linetype = "longdash") + 
  geom_vline(xintercept = as_hms("20:00:00"), colour="red", linetype = "longdash")



```



Plot of peak/off peak energy usage (stacked bar chart?)
- Same plot with shifted load
- Split by months
